Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-14 main.tex
Progress

@@

@@

Informally, a (micro-){\em policy}
is an enriched
step function that works on pairs of contours and machines---i.e., it is a
partial function taking a contour and a machine state to an updated contour
and machine state that, wherever is is defined, agrees with
$\stepstounderfull$.  That is,
it represents an enforcement mechanism that sometimes halts programs earlier
than the bare step relation does, but that agrees with the bare step
relation when it allows evaluation to proceed.  We write
$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.

Formally, a policy
%
$\pol \in \CONTOURS \rightarrow \exists X.\; X \times (\MACHS \times X
\rightharpoonup \MACHS \times X)$
%
 is a function that takes a contour to
\begin{enumerate*}
  \item a private type $X$,
  \item an initial value of $X$, and
  \item a partial step function of type
  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
  agrees with $\stepstounderfull$ on machine states.
\end{enumerate*}
That is, a policy takes a starting contour and produces a runtime monitor
that carries along some private information and uses it to decide whether to
terminate execution early.

\bcp{Here is where it starts getting iffy...}

A {\em contoured (hyper)property} is a function that takes a contour to a
hyperproperty.  That is, it takes a starting contour and produces a
\bcp{prefix-closed?} set of pairs of sequences of machine states.

The {\em observation trace} arising from a finite or infinite sequence of
contours and machine states
$(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
$\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$ is a
partial map taking each machine component to the (finite or infinite)
sequence of {\em distinct} values taken on by this component at moments when
it is {\em observable}.  Formally, this map can be constructed as
follows. For each component $\component$,
%
(1) if
$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
$\obstrace(\component) = \bot$;
%
and otherwise (2), $\obstrace(\component)$ is a sequence whose first element
is $\mach_i(\component)$ for the smallest $i$ such that
$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
than the index $i$ of the $m^{th}$ element and such that
$\contour_j(\component) = (\underscore,\OBS)$ and
$\mach_j(\component) \neq \mach_i(\component)$.
%
\iftext
\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
\newcommand{\aux}{\mathit{aux}}
\bcp{Don't like the ``rest'' argument and the fact that we don't
  have a metavariable for sequences of pairs of contours and machines.
  Consider defining one.}
%
\[
\begin{array}{lcl}
\aux \; ( [\,] ) \; \trace& = & \trace
\\
\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
  & = &
        \aux \;( \mathit{rest} )
        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
\\
\\
\obstrace \;( (\contour_0,\mach_0),\, \ldots )
  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
      \component.\, \bot)
\end{array}
\]
\fi

@@

\bcp{Recheck.}
Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\pol$.

@@

\subsection{More...}
  \begin{itemize}
  \item Proposal: Maybe forget about full-blown NI and focus the whole paper
  on variants of ``lockstep NI'' where the PC is the same in every pair of
  related runs?
  \item One issue that I am not sure about: Could there be a sneaky way of
  ``putting secret information into the control state'' where the PCs of
  related runs are always in lockstep but, because the instruction memory is
  writeable, different instructions can get dispatched.

@@

  \item One nice observation is that, if we are only interested in pairs of
  runs with identical PCs, there are two different ways to proceed: (1) say
  that we {\em ignore} runs where the PCs (or instruction streams) diverge
  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
  instruction streams) diverge, i.e., demand that this never happens (e.g.,
  stack and heap policies).
  \item A remaining question is what to do about the possibility that the
  contour changes differently on two related runs.  We could disallow this,
  but not certain it's necessary---maybe everything works out OK.
  \end{itemize}

>>>>

A {\em contoured trace} is a finite or infinite sequence
$\seq{(\contour,\mach)} = (\contour_0,\mach_0),\,
(\contour_1,\mach_1),\ldots$ of (contour, machine state) pairs.

@@

\subsection{Observations}

A {\em component observation} is a finite or infinite sequence of words in
which adjacent elements are always distinct (``no stuttering'').  An {\em
  observation} is a map from machine components to component observations.

The observation arising from a contoured trace $\seq{(\contour,\mach)}$,
written $\obstrace(\seq{(\contour,\mach)})$, is a map taking each machine
component to the (finite or infinite) sequence of the {\em distinct} values
taken on by this component at moments when it is {\em observable}.

Formally, $\obstrace(\seq{(\contour,\mach)})$ can be constructed as
follows. For each component $\component$,
%
if
$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
$\obstrace(\component) = \bot$;
%
otherwise, $\obstrace(\component)$ is a sequence whose first element
is $\mach_i(\component)$ for the smallest $i$ such that
$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
than the index $i$ of the $m^{th}$ element and such that
$\contour_j(\component) = (\underscore,\OBS)$ and
$\mach_j(\component) \neq \mach_i(\component)$.
%
\iftext
\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
\newcommand{\aux}{\mathit{aux}}
\bcp{Don't like the ``rest'' argument and the fact that we don't
  have a metavariable for sequences of (contour, machine state) pairs.
  Consider defining one.}
%
\[
\begin{array}{lcl}
\aux \; ( [\,] ) \; \trace& = & \trace
\\
\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
  & = &
        \aux \;( \mathit{rest} )
        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
\\
\\
\obstrace \;( (\contour_0,\mach_0),\, \ldots )
  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
      \component.\, \bot)
\end{array}
\]
\fi

% A {\em contoured (hyper)property} is a function that takes a contour to a
% hyperproperty.  That is, it takes a starting contour and produces a
% \bcp{prefix-closed?} set of pairs of sequences of machine states.

% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
% \times X) $
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a function that allows ``reading off'' a contour from an $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup X$.
% \end{enumerate*}

A {\em contoured hyperproperty} is a set of pairs of contoured
traces\iftext{} \bcp{Maybe prefix closed?}\fi.
\bcp{Should we require that it contain all possible starting states and
  contours (though of course the sequences can be singletons in many
  cases!)?  What does it mean if it doesn't?}

@@

Informally, a (micro-){\em policy} is an enriched step function that,
whenever is is defined, agrees with $\stepstounderfull$.  It represents an
enforcement mechanism that sometimes halts programs earlier than the bare
step relation does but that agrees with the bare step relation as long as it
allows evaluation to proceed.  We write
$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.

Formally, a policy
%
$\pol \in \CONTOURS \rightarrow \MACHS \rightarrow \SEQ{\MACHS}$
%
takes a contour $\contour$ and a starting machine state $\mach_0$ and produces
a finite or infinite sequence of states $\mach_0,\,\mach_1 ,\, \ldots$ such
that $\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the sequence.

% \bcp{Fix: even simpler and cleaner, just run the step function to completion
%   and forget about $X$---i.e., a property is just a function from a pair of
%   a contour and a machine state to a finite or infinite sequence of machine
%   states!}
% Formally, a policy
% %
% $\pol \in \exists X.\; (\CONTOURS \rightarrow X) \times (\MACHS \times X
% \rightharpoonup \MACHS \times X)$
% %
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
%   agrees with $\stepstounderfull$ on machine states. \bcp{We could actually
%     just make it return an X and calculate the machine state from the
%     ordinary step function.  This is perhaps a bit cleaner.}
% \end{enumerate*}
% That is, a policy takes a starting contour and produces a runtime monitor
% that carries along some private information and uses it to decide whether to
% terminate execution early.

Policy $\pol$ {\em guarantees} property $\prop$ if \bcp{......}

% every $\contour$,
% every pair $(\mach,\mach')$ of variant machine states with respect to
% $\contour$ are indistinguishable under $\pol$.

@@

% Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
% every pair $(\mach,\mach')$ of variant machine states with respect to
% $\contour$ are indistinguishable under $\pol$.

@@

\iftext
\subsection{More...}
  \begin{itemize}
  \item Proposal: Maybe forget about full-blown NI and focus the whole paper
  on variants of ``lockstep NI'' where the PC is the same in every pair of
  related runs?
  \item One issue that I am not sure about: Could there be a sneaky way of
  ``putting secret information into the control state'' where the PCs of
  related runs are always in lockstep but, because the instruction memory is
  writeable, different instructions can get dispatched.

@@

  \bcp{Answer: Just make instructions read-only.}

  \item One nice observation is that, if we are only interested in pairs of
  runs with identical PCs, there are two different ways to proceed: (1) say
  that we {\em ignore} runs where the PCs (or instruction streams) diverge
  (e.g., taint tracking), or (2) {\em disallow} runs where the PCs (or
  instruction streams) diverge, i.e., demand that this never happens (e.g.,
  stack and heap policies).
  \item A remaining question is what to do about the possibility that the
  contour changes differently on two related runs.  We could disallow this,
  but not certain it's necessary---maybe everything works out OK.
  \end{itemize}
\fi
